Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing

### 2020-02-06 Dissonance_Santiago_v2/Dissonance_Santiago_v2.tex
Reorganized folder structure and created a new version to revise the paper after receiving reviewer comments from Journal of Transportation and Health.

  \journal{Journal of Transportation \& Health} % Sets Journal name

@@

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)
% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

  \title{Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance
from the Perspective of Subjective Wellbeing}
    \author[Some University]{Author 1\corref{c1}}
   \ead{a1@university.com} 
   \cortext[c1]{Corresponding Author}
    \author[Some School]{Author 2}
   \ead{a2@school.com}

      \address[Some University]{Department, Street, City, State, Zip}
    \address[Some School]{Department, Street, City, State, Zip}

  \begin{abstract}
  \textbf{Introduction}\\
  Subjective wellbeing is a topic that has attracted considerable
  attention in recent years due to the way it correlates with health. From
  a transportation perspective, there is a burgeoning literature on the
  way travel can impact subjective wellbeing, and how this, in turn, can
  influence behavior.\\
  \textbf{Objective}\\
  The objective of this paper is to analyze a number of affective values
  associated with subjective wellbeing and the modes of transportation
  that people commonly use. In particular, we are interested in the
  potential for dissonance with respect to the primary mode of travel.\\
  \textbf{Materials and Methods}\\
  The study is based on data collected from a sample of travellers in the
  city of Santiago, in Chile. Participants in the study were asked about
  their usual mode of travel, and then were asked to name the mode or
  modes that they associate with the affective values of freedom,
  enjoyment, happiness, poverty, luxury and status. Analysis is based on
  tests of independence and visualization via mosaic plots.\\
  \textbf{Results}\\
  The results indicate that users of public transportation experience the
  most dissonance in terms of affective values, and active travellers the
  least. For those travellers who experience dissonance, active travel is
  the mode most commonly associated with freedom, enjoyment, and
  happiness, public transportation is most commonly associated with
  poverty, and the automobil is most commonly associated with luxury and
  status.
  \end{abstract}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/plot-descriptive-statistics-1.pdf}
\caption{\label{fig:descriptive-statistics}Descriptive statistics of the
sample}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-primary-mode-travel-1.pdf}
\caption{\label{fig:primary-mode-travel}Frequency of primary mode used
for regular commute; top panel: all modes, bottom panel: aggregated
modes}
\end{figure}

@@

The preceding analysis suggests that there is significant mode
dissonance along various dimensions and for various affective values.
This is for the most part in line with previous research, although by
examining different affective values individually instead of a summary
measure of wellbeing, we are able to differentiate better in terms of
how travellers respond to different affects. Less is known about the
values that travellers associate with modes \emph{other} than the one
they use. For this reason, after developing a profile of the travelers
who experience mode dissonance in the preceding section, we are now
interested in the responses of travellers with respect to the modes they
tend to associate with various affective values. For this analysis we
employ a visualization technique known as a mosaic plot. Mosaic plots
are used in the exploration of multivariate categorical data (Friendly,
1994; Hofmann, 2000), and they provide visual representations of
multi-way contingency tables. The elements of the mosaic plot are called
\emph{tiles}, and the dimensions of each tile are proportional to values
of variables in the underlying table.

We begin our analysis by plotting primary mode of travel and the modes
associated with each affective value (see Figure
\ref{fig:mosaic-plots-by-attribute}). The height of tile \(j\) in these
mosaics corresponds to the proportion of travellers who primarily travel
by mode \(j\), hence we can see that the most common primary mode of
travel in this sample is by public transportation, followed by car,
active travel, and other (also see Figure
\ref{fig:primary-mode-travel}). Notice that there are some small
discrepancies in the heights across affective values: these are caused
by a small number of non-responses. Next, the width of the tile \(k\) in
the mosaics is proportional to the frequency with which a mode was named
in relation to affective value \(i\). Recall that respondents could
nominate more than one mode in response to each affective value, and
some did. For this reason, we used sampling weights, so that the total
number of responses corresponds to the size of the sample.

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-1.pdf}
\caption{\label{fig:mosaic-plots-by-attribute}Mosaic plots for affective
values; the y-axis is the proportion of respondents by main mode of
transportation, and the x-axis is the proportion of modes selected for
each value}
\end{figure}

@@

There are some interesting differences when viewed from the perspective
of travellers of different ages (see Figure
\ref{fig:mosaic-plots-by-age}). For example, as seen above, active
travel is commonly associated with freedom, happiness, and enjoyment,
even by car users, but especially by users of public transportation and
other modes. However, when we break this down by age, we notice that
this tendency weakens as people age, and older travellers increasingly
associate these affective values to the car. Furthermore, the tendency
to associate status with the car tends to increase with age, with the
possible exception of public transport users, for whom other options
(possibly taxi), gain status for older travellers (\textgreater{} 54) at
the expense of other modes, including the car. In contrast, the car
loses in the value of luxury with age, while other options become more
commonly associated with feelings of luxury for older respondents, even
for car users (Figure 2). Older car users (\textgreater{} 54) show less
dissonance in status than in luxury, while active travellers present
more dissonance in status than in luxury - and in fact, they more
commonly project status in the car.

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-age-1.pdf}
\caption{\label{fig:mosaic-plots-by-age}Mosaic plots for affective
values by age; the y-axis is the proportion of respondents by main mode
of transportation, and the x-axis is the proportion of modes selected
for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared
Test)}
\end{figure}

@@

Only two affective values were significant in the preceding analysis
when seen from the perspective of education, namely luxury and status.
As seen in Figure \ref{fig:mosaic-plots-by-education}, the perceptions
of luxury and status are fairly consistent when differentiated by
education. Most respondents tend to associate luxury and status to the
use of car, albeit this association is stronger for status than for
luxury. The second mode most often mentioned in relation to luxury is
``other'' likely due to the flexibility and convenience (but cost) of
using taxis. The exception in this case is the traveller with graduate
education who uses active modes. This type of respondent, while still
more likely to associate luxury and status with the car, more commonly
links active travel to these affects than travellers with lower levels
of education. Most public transport users tend to associate luxury and
status to the car, however, they associate active modes with status more
than the same group with respect to luxury.

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-education-1.pdf}
\caption{\label{fig:mosaic-plots-by-education}Mosaic plots for affective
values by education; the y-axis is the proportion of respondents by main
mode of transportation, and the x-axis is the proportion of modes
selected for each value (p-values are for Cochran-Mantel-Haenszel
Chi-Squared Test)}
\end{figure}

@@

Three affective values were significant in the analysis by income;
perhaps not surprisingly, all three have an economic dimension. Figure
\ref{fig:mosaic-plots-by-income} reflects some important social
inequalities associated with the way users feel about transportation in
Santiago. When the affective perception of poverty is analyzed, the
image shows that generally public transport is the mode most commonly
associated with poverty. There are some interesting differences,
however, to be observed by income level. Whereas middle income users of
public transportation associate this mode to poverty most frequently,
this is not the case of high income users of public transportation, who
tend to link active travel to poverty with greater frequency. Similarly,
active travel is more often associated with poverty by low income
travellers than middle or high income travellers. A possible explanation
for this is that walking or cycling are related to the idea of a lack of
financial means to access other modes of transport. For high income car
users, on the other hand, public transportation may have connotations of
poverty due to its massive use, whereas active travel may be seen more
as a lifestyle alternative.

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-income-1.pdf}
\caption{\label{fig:mosaic-plots-by-income}Mosaic plots for affective
values by income; the y-axis is the proportion of respondents by main
mode of transportation, and the x-axis is the proportion of modes
selected for each value (p-values are for Cochran-Mantel-Haenszel
Chi-Squared Test)}
\end{figure}

@@

The last dimension that we examine is typical commute time. Four
affective values were significant along this dimension in Table
\ref{tab:cross-tabulation-results}. Figure
\ref{fig:mosaic-plots-by-travel-time} shows the way travellers associate
different affective values to modes of transportation by length of
typical commute. The first couple of factors, freedom and enjoyment, are
associated with eudaimonic and hedonic factors of wellbeing, as
discussed earlier in the paper. On the other hand, the luxury and status
are defined mainly by a socio-economic component. As seen in the figure,
freedom and enjoyment are most frequently mentioned in relation to
active travel, followed by car. However, active travel is more
frequently associated with freedom and enjoyment by people who typically
undertake relatively short trips (\textless20 min) or travellers whose
typical trips are longer (\textgreater60 min). In this respect, active
travel may be seen as an ideal mode for trips, and as an enviable mode
by people whose trips are longer.

In the case of luxury and status factors, both follow a similar pattern,
despite variations in travel times. Despite the travel time exceeding 60
minutes, car users consider that their transport mode is mostly
associated with luxury and status. The users of active transport in the
sample tend to travel less than 20 minutes, and this is the same segment
that considers that the status and luxury factors correspond mainly to
the car and then to active transport. Looking at the four factors set
out in Figure \ref{fig:mosaic-plots-by-travel-time}, it could be
understood that the modes that are associated with intrinsically
positive characteristics such as freedom and enjoyment tend to use
public transport. On the contrary, factors such as luxury and status
tend to have a negative connotation, where the association with the car
and other modes of transport (such as the individual taxi and the
collective taxi) tends to dominate.

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v1_files/figure-latex/figure-mosaic-plots-by-attribute-and-time-1.pdf}
\caption{\label{fig:mosaic-plots-by-travel-time}Mosaic plots for
affective values by commute time; the y-axis is the proportion of
respondents by main mode of transportation, and the x-axis is the
proportion of modes selected for each value (p-values are for
Cochran-Mantel-Haenszel Chi-Squared Test)}
\end{figure}

>>>>

@@

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header
\usepackage[margin=1in]{geometry}
\usepackage{lineno}
\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

  \title{Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance
from the Perspective of Subjective Wellbeing}
    \author[Some University]{Author 1\corref{Corresponding Author}}
   \ead{a1@university.com} 
    \author[Some School]{Author 2}
   \ead{a2@school.com} 
      \address[Some University]{Department, Street, City, State, Zip}
    \address[Some School]{Department, Street, City, State, Zip}

  \begin{abstract}
  \textbf{Introduction}\\
  Subjective wellbeing is a topic that has attracted considerable
  attention in recent years due to the way it correlates with health. From
  a transportation perspective, there is a burgeoning literature on the
  way travel can impact subjective wellbeing, and how this, in turn, can
  influence behavior.\\
  \textbf{Objective}\\
  The objective of this paper is to analyze a number of affective values
  associated with subjective wellbeing and the modes of transportation
  that people commonly use. In particular, we are interested in the
  potential for dissonance with respect to the primary mode of travel.\\
  \textbf{Materials and Methods}\\
  The study is based on data collected from a sample of travellers in the
  city of Santiago, in Chile. Participants in the study were asked about
  their usual mode of travel, and then were asked to name the mode or
  modes that they associate with the affective values of freedom,
  enjoyment, happiness, poverty, luxury and status. Analysis is based on
  tests of independence and visualization via mosaic plots.\\
  \textbf{Results}\\
  The results indicate that users of public transportation experience the
  most dissonance in terms of affective values, and active travellers the
  least. For those travellers who experience dissonance, active travel is
  the mode most commonly associated with freedom, enjoyment, and
  happiness, public transportation is most commonly associated with
  poverty, and the automobil is most commonly associated with luxury and
  status.

  journal: ``Journal of Transportation \& Health''
  \end{abstract}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/plot-descriptive-statistics-1.pdf}
\caption{\label{fig:descriptive-statistics}Descriptive statistics of the
sample}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-primary-mode-travel-1.pdf}
\caption{\label{fig:primary-mode-travel}Frequency of primary mode used
for regular commute; top panel: all modes, bottom panel: aggregated
modes}
\end{figure}

@@

The preceding analysis suggests that there is significant mode
dissonance along various dimensions and for various affective values.
This is for the most part in line with previous research, although by
examining different affective values individually instead of a summary
measure of wellbeing, we are able to differentiate better in terms of
how travellers respond to different affects. Less is known about the
values that travellers associate with modes \emph{other} than the one
they use. For this reason, after developing a profile of the travelers
who experience mode dissonance in the preceding section, we are now
interested in the responses of travellers with respect to the modes they
tend to associate with various affective values.

\textbf{\textbf{For this analysis we employ a visualization technique
known as a mosaic plot. Mosaic plots are used in the exploration of
multivariate categorical data (Friendly, 1994; Hofmann, 2000), and they
provide visual representations of multi-way contingency tables. The
elements of the mosaic plot are called \emph{tiles}, and the dimensions
of each tile are proportional to values of variables in the underlying
table.}}

We begin our analysis by plotting primary mode of travel and the modes
associated with each affective value (see Figure
\ref{fig:bar-plots-by-attribute}). The height of tile \(j\) in these
mosaics corresponds to the proportion of travellers who primarily travel
by mode \(j\), hence we can see that the most common primary mode of
travel in this sample is by public transportation, followed by car,
active travel, and other (also see Figure
\ref{fig:primary-mode-travel}). Notice that there are some small
discrepancies in the heights across affective values: these are caused
by a small number of non-responses. Next, the width of the tile \(k\) in
the mosaics is proportional to the frequency with which a mode was named
in relation to affective value \(i\). Recall that respondents could
nominate more than one mode in response to each affective value, and
some did. For this reason, we used sampling weights, so that the total
number of responses corresponds to the size of the sample.

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-bar-plots-by-attribute-1.pdf}
\caption{\label{fig:bar-plots-by-attribute}Plots for affective values;
in the y-axis are the number of cases by primary mode of transportation,
and in the x-axis are the modes selected for each affective value}
\end{figure}

@@

There are some interesting differences when viewed from the perspective
of travellers of different ages (see Figure \ref{fig:bar-plots-by-age}).
For example, as seen above, active travel is commonly associated with
freedom, happiness, and enjoyment, even by car users, but especially by
users of public transportation and other modes. However, when we break
this down by age, we notice that this tendency weakens as people age,
and older travellers increasingly associate these affective values to
the car. Furthermore, the tendency to associate status with the car
tends to increase with age, with the possible exception of public
transport users, for whom other options (possibly taxi), gain status for
older travellers (\textgreater{} 54) at the expense of other modes,
including the car. In contrast, the car loses in the value of luxury
with age, while other options become more commonly associated with
feelings of luxury for older respondents, even for car users (Figure 2).
Older car users (\textgreater{} 54) show less dissonance in status than
in luxury, while active travellers present more dissonance in status
than in luxury - and in fact, they more commonly project status in the
car.

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-bar-plots-by-attribute-and-age-1.pdf}
\caption{\label{fig:bar-plots-by-age}Plots for affective values by age;
in the y-axis is the proportion in the interval {[}0, 1{]}, and the
x-axis is the mode selected for each value (p-values are for
Cochran-Mantel-Haenszel Chi-Squared Test)}
\end{figure}

@@

\textbf{\textbf{Only two affective values were significant in the
preceding analysis when seen from the perspective of education, namely
luxury and status. As seen in Figure \ref{fig:bar-plots-by-education},
the perceptions of luxury and status are fairly consistent when
differentiated by education. Most respondents tend to associate luxury
and status to the use of car, albeit this association is stronger for
status than for luxury. The second mode most often mentioned in relation
to luxury is ``other'' likely due to the flexibility and convenience
(but cost) of using taxis. The exception in this case is the traveller
with graduate education who uses active modes. This type of respondent,
while still more likely to associate luxury and status with the car,
more commonly links active travel to these affects than travellers with
lower levels of education. Most public transport users tend to associate
luxury and status to the car, however, they associate active modes with
status more than the same group with respect to luxury.}}

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-bar-plots-by-attribute-and-education-1.pdf}
\caption{\label{fig:bar-plots-by-education}Plots for affective values by
level of education; in the y-axis is the proportion in the interval
{[}0, 1{]}, and the x-axis is the mode selected for each value (p-values
are for Cochran-Mantel-Haenszel Chi-Squared Test)}
\end{figure}

@@

Three affective values were significant in the analysis by income;
perhaps not surprisingly, all three have an economic dimension. Figure
\ref{fig:bar-plots-by-income} reflects some important social
inequalities associated with the way users feel about transportation in
Santiago. When the affective perception of poverty is analyzed, the
image shows that generally public transport is the mode most commonly
associated with poverty. There are some interesting differences,
however, to be observed by income level. Whereas middle income users of
public transportation associate this mode to poverty most frequently,
this is not the case of high income users of public transportation, who
tend to link active travel to poverty with greater frequency. Similarly,
active travel is more often associated with poverty by low income
travellers than middle or high income travellers. A possible explanation
for this is that walking or cycling are related to the idea of a lack of
financial means to access other modes of transport. For high income car
users, on the other hand, public transportation may have connotations of
poverty due to its massive use, whereas active travel may be seen more
as a lifestyle alternative.

@@

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-bar-plots-by-attribute-and-income-1.pdf}
\caption{\label{fig:bar-plots-by-income}Plots for affective values by
level of income; in the y-axis is the proportion in the interval {[}0,
1{]}, and the x-axis is the mode selected for each value (p-values are
for Cochran-Mantel-Haenszel Chi-Squared Test)}
\end{figure}

@@

The last dimension that we examine is typical commute time. Four
affective values were significant along this dimension in Table
\ref{tab:cross-tabulation-results}. Figure
\ref{fig:bar-plots-by-travel-time} shows the way travellers associate
different affective values to modes of transportation by length of
typical commute. The first couple of factors, freedom and enjoyment, are
associated with eudaimonic and hedonic factors of wellbeing, as
discussed earlier in the paper. On the other hand, the luxury and status
are defined mainly by a socio-economic component. As seen in the figure,
freedom and enjoyment are most frequently mentioned in relation to
active travel, followed by car. However, active travel is more
frequently associated with freedom and enjoyment by people who typically
undertake relatively short trips (\textless20 min) or travellers whose
typical trips are longer (\textgreater60 min). In this respect, active
travel may be seen as an ideal mode for trips, and as an enviable mode
by people whose trips are longer.

In the case of luxury and status factors, both follow a similar pattern,
despite variations in travel times. Despite the travel time exceeding 60
minutes, car users consider that their transport mode is mostly
associated with luxury and status. The users of active transport in the
sample tend to travel less than 20 minutes, and this is the same segment
that considers that the status and luxury factors correspond mainly to
the car and then to active transport. Looking at the four factors set
out in Figure \ref{fig:bar-plots-by-travel-time}, it could be understood
that the modes that are associated with intrinsically positive
characteristics such as freedom and enjoyment tend to use public
transport. On the contrary, factors such as luxury and status tend to
have a negative connotation, where the association with the car and
other modes of transport (such as the individual taxi and the collective
taxi) tends to dominate.

\begin{figure}
\centering
\includegraphics{Dissonance_Santiago_v2_files/figure-latex/figure-bar-plots-by-attribute-and-travel-time-1.pdf}
\caption{\label{fig:bar-plots-by-travel-time}Plots for affective values
by commute time; in the y-axis is the proportion in the interval {[}0,
1{]}, and the x-axis is the mode selected for each value (p-values are
for Cochran-Mantel-Haenszel Chi-Squared Test)}
\end{figure}
